---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(lme4)
library(psych)
library(skimr)
library(tidymodels)


test <- read_csv('~/Kaggle/titanic/CSVs/test.csv')

train <- read_csv('~/Kaggle/titanic/CSVs/train.csv')

gender_submission <- read_csv('~/Kaggle/titanic/CSVs/gender_submission.csv')

```

```{r}
psych::describe(gender_submission)

skimr::skim(gender_submission)
```

# Data cleaning

Variable Definition Key

survival: Survival 0 = No, 1 = Yes

pclass Ticket class 1 = 1st, 2 = 2nd, 3 = 3rd

sex: Sex\
Age: Age in years\
sibsp \# of siblings / spouses aboard the Titanic\
parch \# of parents / children aboard the Titanic\
ticket Ticket number\
fare Passenger fare\
cabin Cabin number\
embarked Port of Embarkation C = Cherbourg, Q = Queenstown, S = Southampton

```{r}

train <-
  train %>%
  mutate(`Passenger Survived` = factor(Survived, levels = c(0,1), labels = c('No','Yes')),
         `Ticket Class` = factor(Pclass, levels = c(1:3), labels = c('1st','2nd','3rd')),
         `Sex` = as_factor(Sex),
         AgeUnder15 = (Age < 15),
         AgeUnder10 = (Age < 10),
         AgeOver60 = (Age > 60),
         HadParentsOrKidsOnboard = (Parch > 0),
         ParentsOrKidsOnboard = ifelse(Parch == 0, 
                                     'No', 
                                     ifelse(Parch == 1, 
                                            'One', 
                                            'Two or More')),
         HadSibOrSpouseAboard = (SibSp > 0),
         SibsOrSpouseAboard = ifelse(SibSp == 0, 
                                     'No', 
                                     ifelse(SibSp == 1, 
                                            'One', 
                                            'Two or More')),
         logFare = log(Fare))
         

test <-
  test %>%
mutate(`Ticket Class` = factor(Pclass, levels = c(1:3), labels = c('1st','2nd','3rd')),
         `Sex` = as_factor(Sex),
         AgeUnder15 = (Age < 15),
         AgeUnder10 = (Age < 10),
         AgeOver60 = (Age > 60),
         HadParentsOrKidsOnboard = (Parch > 0),
         ParentsOrKidsOnboard = ifelse(Parch == 0, 
                                     'No', 
                                     ifelse(Parch == 1, 
                                            'One', 
                                            'Two or More')),
         HadSibOrSpouseAboard = (SibSp > 0),
         SibsOrSpouseAboard = ifelse(SibSp == 0, 
                                     'No', 
                                     ifelse(SibSp == 1, 
                                            'One', 
                                            'Two or More')),
         logFare = log(Fare))
         
                                 

```

```{r}

glimpse(train)

```

# Exploratory data analysis

## How many survivors?

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`))+
  geom_bar(position = position_dodge2())


```

## Ticket Class

Seems like people in higher classes ought to be more likely to survive...

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = `Ticket Class`)) +
  geom_bar(position = position_dodge2())


```

## Sex

As I recall, women were prioritized for boats...

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = `Sex`)) +
  geom_bar(position = position_dodge2())


```

## Age

Putting in something to capture being a little kid might be helpful...

### Hist

```{r}
train %>%
  ggplot(mapping = aes(x = Age)) +
  geom_histogram()

```

### Relationship to survival

Kids were also prioritized for boats...

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, y = Age)) +
  geom_boxplot()


```

Doesn't actually seem like the ages of dead vs alive mattered...

Might be better as, like, under 15 vs over or something like that

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = `AgeUnder15`)) +
  geom_bar(position = position_dodge2())


```

Or maybe, like, age 10?

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = `AgeUnder10`)) +
  geom_bar(position = position_dodge2())


```

What about old people eg over 60?

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = `AgeOver60`)) +
  geom_bar(position = position_dodge2())


```

## Sibsp

What about if you had someone onboard who was your peer potentially looking out for you?

```{r}

train %>%
  ggplot(mapping = aes(x = SibSp)) +
  geom_histogram()


```

Most people had no sibs/spouses aboard

200 had 1 (probably a spouse for many?)

Some had 2+

Could try looking at it that way...

Seems like bad news if you had lots of other peers aboard for some reason...maybe the 2+ people were people who worked on on the ship?

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = SibsOrSpouseAboard)) +
  geom_bar(position = position_dodge2())


```

## Parch

```{r}

train %>%
  ggplot(aes(x = Parch)) +
  geom_histogram()
```

If you had kids onboard, maybe you were more likely to survive?

Or likewise if you had parents, that made you a kid, so maybe you were more likely to survive?

Seems like you could just treat this as a true/false variable, true if had parents or kids, false if not.

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, fill = ParentsOrKidsOnboard)) +
  geom_bar(position = position_dodge2())


```

## Fare

If you paid more maybe you were more affluent and so less likely to die

```{r}

train %>%
  ggplot(mapping = aes(x  = Fare)) +
  geom_histogram()

train %>%
  ggplot(mapping = aes(x = logFare)) +
  geom_histogram(binwidth = .5)
```

```{r}

train %>%
  ggplot(mapping = aes(x = `Passenger Survived`, y = log(Fare))) +
  geom_boxplot()


```

## Relationships Among Variables

### Is class redundant with fare?

```{r}
psych::polychoric(train$logFare, train$`Ticket Class`)

```

# Models

## Logistic Regression - Predict survival by ticket class and sex

### Make recipe

```{r}

titanic_recipe <- 
  #recipe(`Passenger Survived` ~ `Ticket Class` + Sex,
  recipe(`Passenger Survived` ~ `Ticket Class` + Sex +  SibsOrSpouseAboard + ParentsOrKidsOnboard,
  
         data = train)
```

### Logistic regression model

```{r}
lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm")

```

### Workflow

```{r}
titanic_wflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(titanic_recipe)
```

### Fit the model

```{r}
titanic_logistic_fit <-
  titanic_wflow %>%
  fit(data = train)


```

```{r}
titanic_logistic_fit %>%
  pull_workflow_fit() %>%
  tidy()

```

### Check accuracy on training data

```{r}
train_prediction_check <- 
  predict(titanic_logistic_fit, train) %>%
  bind_cols(train) %>%
  mutate(`Model Prediction is Correct` = (.pred_class == `Passenger Survived`))

table(train_prediction_check$`Model Prediction is Correct` )

```

m

### Make predictions on the test data

```{r}
test_predict <- 
  predict(titanic_logistic_fit, test) %>%
  bind_cols(test %>% select(PassengerId)) %>%
  mutate(Survived = ifelse(.pred_class == 'No',0,1)) %>%
  select(-.pred_class)

write_csv(x = test_predict, file = "~/Kaggle/titanic/test_predictions.csv")
```

# Tree-based approach
